---
title: "Urban environment & postnatal depression: descriptive statistics"
date: 14th June 2021
author: Tim Cadman
subparagraph: true
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{titlesec}
- \usepackage{caption}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{subcaption}
- \definecolor{customcol}{RGB}{38, 70, 83}

output: 
 bookdown::pdf_document2:
  latex_engine: xelatex
  template: "C:/Users/gmj541/OneDrive - University of Bristol/repos/useful-code-r/code/reports/khl_5.tex"
  fig_caption: yes
  toc_depth: 1
---
```{r settings, echo = FALSE}
options(datashield.progress = FALSE)
options(datashield.progress.clear = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(results = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = NA, tidy=FALSE, fig.pos = 'H')
options(knitr.duplicate.label = "allow")
options(knitr.table.format = "latex")
```

```{r render, eval = FALSE, echo = FALSE}
rmarkdown::render(
  input = "./reports/31-may-21/env_pnd_desc_16_07_21.Rmd", 
  output_file = "env-pnd-desc-16-june-21.pdf", envir = globalenv())
```  

```{r load-packages}
library(GGally)   
library(kableExtra)
library(Cairo)
library(see)
library(tidyverse)
library(here)
library(showtext)
```

```{r get-themes}
source("https://raw.githubusercontent.com/timcadman/useful-code-r/master/code/themes/themes.R")
source("https://raw.githubusercontent.com/timcadman/useful-code-r/master/code/themes/lc-names-neat.R")

hVPlot <- function(data = violin.plotData, var, y_label = NULL, limits = NULL){

if(is.null(limits)){
y_scale <- coord_flip(     
  expand = FALSE,
  clip = "on")

} else if(!is.null(limits)){

y_scale <- y_scale <- coord_flip(     
  ylim = limits,
  expand = FALSE,
  clip = "on")

}

data %>%
filter(variable == var) %>%
ggplot(aes(x = cohort_n, y = value)) +
geom_violinhalf(aes(fill = cohort_neat), alpha = 0.75) + 
geom_boxplot(width=0.10, outlier.size = 0.5) +
facet_wrap(~time, ncol = 2, scales = "free_y") +
y_scale + 
ylab(y_label) +
scale_fill_manual(values = palette_ext) + 
theme_std + 
theme_hv +
theme(
  axis.title.y = element_blank()
  )
}

showtext_auto()
```

```{r get-data}

## ---- Descriptives -----------------------------------------------------------
load(here("data", "exp_desc.RData"))
load(here("data", "out_desc.RData"))

## ---- Correlations between exposures -----------------------------------------
load(here("data", "exp_cor_preg.RData"))
load(here("data", "btp_cor_preg.RData"))

## ---- Anonymised plot data ---------------------------------------------------
load(here("data", "violin_out.RData"))

## ---- Exposure - outcome associations ----------------------------------------
load(here("data", "single_reg.RData"))

```

```{r, tidy-data}
time_labs <- c("Birth to 12 months", "Pregnancy")

## ---- Function to tidy descriptives ------------------------------------------
cleanData = function(x){

out <- x %>%
separate(
  data = .,
  col = variable, 
  sep = "_(?=[^_]*$)", 
  into = c("variable", "time")) %>%
mutate(
  .data = ., 
  time = factor(time, labels = time_labs, ordered = TRUE)) %>%
mutate(
  .data = ., 
  time = fct_relevel(time, rev(time_labs))) %>%
left_join(., names_neat, by = "cohort") %>%  
mutate(
  .data = .,
  cohort_neat = fct_relevel(
    cohort_neat, 
    c("MoBa", "DNBC", "GENR", "ALSPAC", "NINFEA", "EDEN Nancy", "EDEN Poitiers", 
      "INMA Gipuzkoa", "INMA Sabadell", "INMA Valencia"))) %>%
mutate(cohort_neat = factor(cohort_neat, ordered = TRUE))

}

desc_clean <- exposures.desc$continuous %>% cleanData

## ---- Violin plot data -------------------------------------------------------
violin.plotData <- violin_out %>% 
bind_rows(.id = "variable") %>% 
cleanData %>% 
group_by(variable, time, cohort) %>% 
mutate(n = n()) %>% 
mutate(cohort_n = paste0(cohort_neat, "\n", "n = ", n))


## ---- Exposure - outcome plot data -------------------------------------------
#single_reg.plotData <- single_reg.out %>% 
#left_join(., names_neat, by = "cohort") %>%  
#mutate(
#  .data = .,
#  cohort_neat = fct_relevel(
#    cohort_neat, 
#    rev(c("MoBa", "ALSPAC", "NINFEA", "Pooled results")))) %>%
#filter(variable != "intercept") %>%
#mutate(across(est:uppci, ~ exp(.x)))
```

\newpage

# Overview
I've put together (i) descriptive plots for exposures and outcomes, (ii) 
correlations between exposures, and (iii) associations between each exposure
and outcome. I will also look at non-linearity in exposure-outcome associations,
but I need to write a function to do spline modelling in DataSHIELD.

\newpage

# Descriptive statistics for exposures

## NDVI (300m buffer)
```{r ndvi, fig.showtext = TRUE}
hVPlot(var = "ndvi300", y_label = "NDVI", limits = c(0.1, 0.8)) 
```

\newpage

## Distance to nearest green spaces > 5000m^2^
There is a mistake with GEN-R data which they are fixing.
```{r green-dist}
hVPlot(
  data = violin.plotData %>% filter(cohort != "genr"), 
  var = "green_dist", 
  y_label = "Distance (m)", 
  limits = c(-100, 1000)) 
```

\newpage

## Distance to nearest blue spaces > 5000m^2^
Again, GEN-R is excluded because of an error with their data.
```{r blue-dist}
hVPlot(
  data = violin.plotData %>% filter(cohort != "genr"), 
  var = "blue_dist", 
  y_label = "Distance (m)", 
  limits = c(-100, 5000)) 
```

\newpage

## Mean NO~2~ in pregnancy
```{r no2-preg}
hVPlot(
  var = "no2", 
  y_label = "N02 (microgram / m3)", 
  limits = c(0, 90))
```

\newpage

## Mean PM~2.5~ in pregnancy  

```{r pm25-preg}
hVPlot(
  var = "pm25", 
  y_label = "N02 (microgram / m3)", 
  limits = c(0, 40))
```

\newpage

## Mean PM~2.5~ from 0-1 years
For some reason this is missing in the data - will check.

## Mean NO~2~ from 0-1 years
Again this is missing but I'm not sure why.

## Mean day-evening-night traffic noise
Missing in INMA Gipuzkoa & Valencia - is this correct?
```{r lden}
hVPlot(
  var = "lden", 
  y_label = "Noise (dBA)", 
  limits = c(30, 80))
```

For ALSPAC (and DNBC?) only the categorical version is available - I will add this.

\newpage

## Mean buildling density
```{r lu-building}
hVPlot(
  var = "bdens300", 
  y_label = "Density (m2 / km2)", 
  limits = c(0, 800000))
```

\newpage

## Percentage of urban green land use
```{r lu-green}
hVPlot(
  var = "urbgr", 
  y_label = "Percent", 
  limits = c(0, 100))
```

\newpage

## Percentage of forest land use
```{r lu-forest}
hVPlot(
  var = "natgr", 
  y_label = "Percent", 
  limits = c(0, 100))
```

\newpage

## Percentage of agricultural land use
```{r lu-agr}
hVPlot(
  var = "agrgr", 
  y_label = "Percent", 
  limits = c(0, 100))
```

\newpage

## Walkability index
```{r lu-walk}
hVPlot(
  var = "walkability_mean", 
  y_label = "Mean", 
  limits = c(0, 0.8))
```

\newpage

## Land use (Shannon's evenness index)
```{r lu-shannon}
hVPlot(
  var = "landuseshan300", 
  y_label = "Eveness index", 
  limits = c(0, 1))
```

\newpage

## Facilities richness
```{r lu-fac-rich}
hVPlot(
  var = "frichness300", 
  y_label = "N facility types / max no. facility types", 
  limits = c(0, 0.6))
```

\newpage

## Facilities density
```{r lu-fac-dens}
hVPlot(
  var = "fdensity300", 
  y_label = "No. facilities within 300m buffer", 
  limits = c(0, 1000))
```

\newpage

# Correlations between exposures

To avoid overload I have just plotted correlations between the pregnancy 
exposures.
```{r}
dh.getCorr <- function(object, remove = c("_preg", "_0_12")){

  var_names <- dimnames(object[[1]])[[2]]
  cor_mat <- object[["Correlation Matrix"]] %>%
  as_tibble() %>%
  purrr::set_names(str_remove(var_names, remove)) %>%
  dplyr::rename(
    lu_green = urbgr, 
    lu_forest = natgr,
    lu_agr = agrgr,
    walkability = walkability_mean)

  return(cor_mat)

  }

coh_names <- c("alspac", "dnbc", "eden_nan", "eden_poit", "inma_gip", "inma_sab", "inma_val", "genr", "moba", "ninfea")

cor_preg.plotdata <- exp_cor_preg %>% 
map(~dh.getCorr(object = ., remove = "_preg")) %>% 
set_names(coh_names)
```

## ALSPAC
```{r, out.width = "100%"}
cor_preg.plotdata$alspac %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_ext[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## DNBC
```{r, out.width = "100%"}
cor_preg.plotdata$dnbc %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_ext[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## EDEN Poitieres
```{r}
cor_preg.plotdata$eden_poit %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_ext[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## EDEN Nancy
```{r}
cor_preg.plotdata$eden_nan %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_ext[2], mid = "white", high = palette_std[5]) +
theme_cor
```

## INMA Gipuzkoa
```{r}
cor_preg.plotdata$inma_gip %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## INMA Sabadell
```{r}
cor_preg.plotdata$inma_sab %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## INMA Valencia
```{r}
cor_preg.plotdata$inma_val %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## GEN-R
```{r}
cor_preg.plotdata$genr %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## MoBa
```{r}
cor_preg.plotdata$moba %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## NINFEA
```{r}
cor_preg.plotdata$ninfea %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage


# Correlations between timepoints

Note I still need to get access to INMA natural space variables age 0-1.
```{r}
tidyBtpCor <- function(x){
x %>%
map(~.[["Correlation Matrix"]][[2]]) %>%
setNames(coh_names) %>%
bind_rows()
}

btp_cor %>% map(tidyBtpCor) %>%
setNames(
  c(
  "lden", "ndvi300", "green_dist", "blue_dist", "bdens300", "fdensity300", 
  "frichness300", "landuseshan300", "walkability_mean", "agrgr", "natgr", 
    "urbgr")) %>%
bind_rows(.id = "variable") 
```



# Descriptive statistics for outcome
Here are the frequencies for postnatal depression. We need discuss how to code
this for DNBC.
```{r}
out_clean <- outcome.desc$categorical %>%
filter(cohort != "inma") %>%
left_join(., names_neat, by = "cohort") %>%
mutate(
  cohort_neat = fct_relevel(
    cohort_neat, 
    c("MoBa", "GENR", "ALSPAC", "NINFEA", "Pooled results"))
  )

out_clean %>%
filter(cohort != "combined") %>%
  ggplot(aes(x = category, y = perc_total, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~cohort, ncol = 10) +
  ylab("Percentage") +
  coord_cartesian(
    ylim = c(0, 100), 
    expand = FALSE) +
  scale_fill_manual(values = palette_ext) + 
  theme_std + 
  theme_hv +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

\newpage

