---
title: "Urban environment & postnatal depression: descriptive statistics"
date: 2nd June 2021
author: Tim Cadman
subparagraph: true
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{titlesec}
- \usepackage{caption}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{subcaption}
- \definecolor{customcol}{RGB}{38, 70, 83}

output: 
 bookdown::pdf_document2:
  latex_engine: xelatex
  template: "/Users/timcadman/OneDrive - University of Bristol/repos/useful-code-r/code/reports/khl_5.tex"
  fig_caption: yes
  toc_depth: 1
---
```{r settings, echo = FALSE}
options(datashield.progress = FALSE)
options(datashield.progress.clear = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(results = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = NA, tidy=FALSE, fig.pos = 'H')
options(knitr.duplicate.label = "allow")
options(knitr.table.format = "latex")
```

```{r}
library(GGally)   
library(showtext)
library(kableExtra)

source("https://raw.githubusercontent.com/timcadman/useful-code-r/master/code/themes/theme-std.R")
source("https://raw.githubusercontent.com/timcadman/useful-code-r/master/code/themes/lc-names-neat.R")

load(here("data", "exp_desc.RData"))
load(here("data", "out_desc.RData"))

load(here("data", "exp_cor_0_12.RData"))
load(here("data", "exp_cor_preg.RData"))

theme_bar <- theme(
  legend.position = "none",
  panel.grid.major.y=element_line(colour="grey"), 
  panel.grid.major.x=element_line(colour="white"),
  axis.title.x=element_blank(),
  axis.ticks.x=element_blank()
)

theme_cor <- theme(
  plot.background = element_rect(fill =scales::alpha("#CCCCCC", 0.3)),  #Background outside of plot
  plot.title = element_text(family = "avenir-book", hjust = 0.5, vjust=0, size=12, face="bold"), #Plot title, thought don't tend to use
  text=element_text(family = "avenir-book", size=9), #General text 
  axis.title.y = element_text(family = "avenir-book", size=14, margin = margin(t = 0, r = 10, b = 0, l = 0)), #Axis labels
  axis.title.x = element_text(family = "avenir-book", size=14, margin = margin(t = 10, r = 0, b = 0, l = 0)),
  axis.text.x = element_text(family = "avenir-book", size=11, margin = margin(t = 4, r=0, b=0, l=0), colour="black"), #Axis text
  axis.text.y = element_text(family = "avenir-book", size=11, margin = margin(t = 0, r=4, b=0, l=0), colour="black"),
  strip.text.x = element_text(family = "avenir-book", size = 8, face = "bold"),
  legend.background= element_rect(fill=scales::alpha("#CCCCCC", 0.03)), #Legend background colour
  legend.title=element_text(family = "avenir-book", size=8, face="bold"), #Legend title
  legend.text=element_text(family = "avenir-book", size=8), #Legend text
  legend.position="right", #Legend position
  legend.direction="vertical", #Legend stacking
  legend.justification = "left", #Left justify legend
  legend.key.width = unit(3, "line"), #Make amount of line displayed in legend longer
  legend.margin=margin(t=0.2, r=0, b=0.2, l=0, unit="cm")) #Margin around legend

showtext_auto()

```

```{r render, eval = FALSE, echo = FALSE}
rmarkdown::render(
  input = "./reports/31-may-21/exposure-outcome-descriptives.Rmd", 
  output_file = "env-pnd-descriptives-june-21.pdf", envir = globalenv())
```


```{r, echo = FALSE}
time_labs <- c("Birth to 12 months", "Pregnancy")

desc_clean <- exposures.desc %>%
  map(
    ~separate(, 
              data = .,
              col = variable, 
              sep = "_(?=[^_]*$)", 
              into = c("variable", "time"))
  ) %>%
  map(
    ~mutate(
      .data = ., 
      time = factor(time, labels = time_labs, ordered = TRUE))
  ) %>%
  map(
    ~mutate(
      .data = ., 
      time = fct_relevel(time, rev(time_labs)))
  ) %>%
  map(
    ~left_join(., names_neat, by = "cohort")
  ) %>%  
  map(
    ~mutate(
      .data = .,
      cohort_neat = fct_relevel(
        cohort_neat, 
        c("MoBa", "GENR", "ALSPAC", "EDEN", "NINFEA", "INMA")))
  )

```

\newpage

# Overview
I've put together some basic plots of each environmental exposure and also the
postnatal depression outcome. I've split the environmental exposures into two time 
periods where available: (i) pregnancy, (ii) birth - 12 months. You'll see from the 
descriptives there is very little differences in values across the two time periods.

I've also shown the the correlations between environmental variables.  

Luise is still working on the DNBC environmental data - she's encountered some issues with
the upload that she's working with Sido to fix.

# Natural spaces

## NDVI (300m buffer)
```{r ndvi, fig.showtext = TRUE}
showtext_auto()

desc_clean$continuous %>%
  filter(variable == "ndvi300" & cohort != "combined") %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("NDVI") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.8)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Distance to nearest green spaces > 5000m^2^
There's definitely something wrong with GEN-R data. Here is the data for the 
cohorts excluding GEN-R:  

```{r green-dist}
desc_clean$continuous %>%
  filter(
    variable == "green_dist" & !cohort %in% c("genr", "combined")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("Distance (m)") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 400)) +
  scale_fill_manual(values = palette_std[c(1, 3:6)]) + 
  theme_std + 
  theme_bar

```

And here is GEN-R data:
```{r green-dist-genr, results = TRUE}
desc_clean$continuous %>%
  filter(variable == "green_dist" & cohort == "genr") %>%
  select(Cohort = cohort, Time = time, Mean = mean, Std.dev = std.dev) %>%
  kbl() %>% 
  kable_styling(full_width = FALSE, position = "left") %>%
  collapse_rows(valign = "top")
```

\newpage

## Distance to nearest blue spaces > 5000m^2^
Again, something not right with GEN-R data...
```{r blue-dist}
desc_clean$continuous %>%
  filter(
    variable == "blue_dist" & !cohort %in% c("genr", "combined")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("Distance (m)") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 2000)) +
  scale_fill_manual(values = palette_std[c(1, 3:6)]) + 
  theme_std + 
  theme_bar
```

```{r blue-dist-genr, results = TRUE}
desc_clean$continuous %>%
  filter(variable == "blue_dist" & cohort == "genr") %>%
  select(Cohort = cohort, Time = time, Mean = mean, Std.dev = std.dev) %>%
  kbl() %>% 
  kable_styling(full_width = FALSE, position = "left") %>%
  collapse_rows(valign = "top")
```

\newpage

# Air pollution
There only appears to be pregnancy data - do you think this is correct?  

## Mean NO~2~ in pregnancy
```{r no2-preg}
desc_clean$continuous %>%
  filter(variable == "no2" & cohort != "combined") %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  ylab("N02 (microgram / m3)") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 80)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Mean PM~2.5~ in pregnancy  

```{r pm25-preg}
desc_clean$continuous %>%
  filter(variable == "pm25" & cohort != "combined") %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  ylab("PM2.5 (microgram / m3)") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

# Road traffic noise
## Mean day-evening-night traffic noise
This isn't available for ALSPAC  

```{r lden}
desc_clean$continuous %>%
  filter(
    variable == "lden" & cohort != "combined") %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("Noise (dBA)") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 80)) +
  scale_fill_manual(values = palette_std[c(1:2, 4:6)]) + 
  theme_std + 
  theme_bar
```

\newpage

# Land use
I'm waiting on the built environment variables from INMA. I had to send an additional request as they weren't included in the original application (they were suggested later by Payam)  

## Mean buildling density
```{r lu-building}
desc_clean$continuous %>%
  filter(variable == "bdens300" & !cohort %in% c("combined", "inma")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("Density (m2 / km2)") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 750000)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Percentage of urban green land use
Big difference between MoBa and other cohorts - does this look like an error?  

```{r lu-green}
desc_clean$continuous %>%
  filter(variable == "natgr" & !cohort %in% c("combined", "inma")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("Percentage") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 50)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Percentage of forest land use
```{r lu-forest}
desc_clean$continuous %>%
  filter(variable == "agrgr" & !cohort %in% c("combined", "inma")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("Percentage") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Walkability index
```{r lu-walk}
desc_clean$continuous %>%
  filter(variable == "walkability_mean" & !cohort %in% c("combined", "inma")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("Decile") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.6)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Land use (Shannon's evenness index)
```{r lu-shannon}
desc_clean$continuous %>%
  filter(variable == "landuseshan300" & !cohort %in% c("combined", "inma")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.6)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Facilities richness
```{r lu-fac-rich}
desc_clean$continuous %>%
  filter(variable == "frichness300" & !cohort %in% c("combined", "inma")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("N facility types / max no. facility types") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.20)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

## Facilities density
```{r lu-fac-dens}
desc_clean$continuous %>%
  filter(variable == "fdensity300" & !cohort %in% c("combined", "inma")) %>%
  ggplot(aes(x = cohort_neat, y = mean, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~time, ncol = 1) +
  ylab("No. facilities within 300m buffer") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.20)) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar
```

\newpage

# Correlations between exposures

To avoid overload I have just plotted correlations between the pregnancy exposures. It's
easy to add the exposures in the first year of life if you want to see them too.  

```{r}

  dh.getCorr <- function(object, remove = c("_preg", "_0_12")){

  var_names <- dimnames(object[[1]])[[2]]
  cor_mat <- object[["Correlation Matrix"]] %>%
  as_tibble() %>%
  purrr::set_names(str_remove(var_names, remove)) %>%
  dplyr::rename(
    lu_green = urbgr, 
    lu_forest = natgr,
    lu_agr = agrgr,
    walkability = walkability_mean)

  return(cor_mat)

  }

coh_names <- c("alspac", "eden", "inma", "genr", "moba", "ninfea")

cor_preg.plotdata <- exp_cor_preg %>% 
map(~dh.getCorr(object = ., remove = "_preg")) %>% 
set_names(coh_names)

#cor_0_12.plotdata <- exp_cor_0_12 %>% 
#map(~dh.getCorr(object = ., remove = "_0_12")) %>% 
#set_names(coh_names)
```

## ALSPAC
```{r, out.width = "100%"}
cor_preg.plotdata$alspac %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## EDEN
```{r}
cor_preg.plotdata$eden %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## INMA
See above - waiting on built environment data.
```{r}
cor_preg.plotdata$inma %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## GEN-R
```{r}
cor_preg.plotdata$genr %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## MoBa
```{r}
cor_preg.plotdata$moba %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

## NINFEA
```{r}
cor_preg.plotdata$ninfea %>%
ggcorr(data = NULL, cor_matrix = ., label = TRUE, label_round = 2, hjust = 1, layout.exp = 1, label_size = 3, low = palette_std[2], mid = "white", high = palette_std[5]) +
theme_cor
```

\newpage

# Postnatal depression
You can see something is clearly miscoded for MoBa as levels of PND can't be this high - I will contact them.  

I am working with Luise to code the DNBC PND variable.  

Values look plausible for the other cohorts.
```{r}
out_clean <- outcome.desc$categorical %>%
left_join(., names_neat, by = "cohort") %>%
mutate(
  cohort_neat = fct_relevel(
    cohort_neat, 
    c("MoBa", "GENR", "ALSPAC", "EDEN", "NINFEA", "INMA"))
  )
  
out_clean %>%
filter(cohort != "combined") %>%
  ggplot(aes(x = category, y = perc_total, fill = cohort_neat)) +
  geom_bar(stat = "identity") +
  facet_wrap(~cohort, ncol = 7) +
  ylab("Percentage") +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100), breaks = c(seq(0, 100, 10))) +
  scale_fill_manual(values = palette_std) + 
  theme_std + 
  theme_bar +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Queries / discussion

## Next steps
1. Test for non-linear associations between each exposure and outcome
2. Contact cohort(s) about issues with data
3. Harmonise DNBC PND variable
4. Descriptives for DNBC exposures when available

## Queries
1. How to include all these exposures without risk of false positives?
2. Harmonisation of DNBC PND variable

