---
title: "Urban environment & postnatal depression: descriptive statistics"
date: 16th July 2021
author: Tim Cadman
subparagraph: true
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{titlesec}
- \usepackage{caption}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage{subcaption}
- \definecolor{customcol}{RGB}{38, 70, 83}

output: 
 bookdown::pdf_document2:
  latex_engine: xelatex
  template: "/Users/timcadman/OneDrive - University of Bristol/repos/useful-code-r/code/reports/khl_5.tex"
  fig_caption: yes
  toc_depth: 1
---
```{r settings, echo = FALSE}
options(datashield.progress = FALSE)
options(datashield.progress.clear = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(results = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = NA, tidy=FALSE, fig.pos = 'H')
options(knitr.duplicate.label = "allow")
options(knitr.table.format = "latex")
```

```{r render, eval = FALSE, echo = FALSE}
rmarkdown::render(
  input = "./reports/31-may-21/test-fig-height.Rmd", 
  output_file = "env-pnd-desc-test-fig-height.pdf", envir = globalenv())
```  

```{r load-packages}
library(GGally)   
library(kableExtra)
library(Cairo)
library(see)
library(tidyverse)
library(here)
library(showtext)
```

```{r get-themes}
source("https://raw.githubusercontent.com/timcadman/useful-code-r/master/code/themes/themes.R")
source("https://raw.githubusercontent.com/timcadman/useful-code-r/master/code/themes/lc-names-neat.R")

hVPlot <- function(data = violin.plotData, var, y_label = NULL, limits = NULL){

if(is.null(limits)){
y_scale <- coord_flip(     
  expand = FALSE,
  clip = "on")

} else if(!is.null(limits)){

y_scale <- y_scale <- coord_flip(     
  ylim = limits,
  expand = FALSE,
  clip = "on")

}

data %>%
filter(variable == var) %>%
ggplot(aes(x = cohort_n, y = value)) +
geom_violinhalf(aes(fill = cohort_neat), alpha = 0.75) + 
geom_boxplot(width=0.10, outlier.size = 0.5) +
facet_wrap(~time, ncol = 2, scales = "free_y") +
y_scale + 
ylab(y_label) +
scale_fill_manual(values = palette_ext) + 
theme_std + 
theme_hv +
theme(
  axis.title.y = element_blank()
  )
}

showtext_auto()
```

```{r get-data}

## ---- Descriptives -----------------------------------------------------------
load(here("data", "exp_desc.RData"))
load(here("data", "out_desc.RData"))

## ---- Correlations between exposures -----------------------------------------
load(here("data", "exp_cor_preg.RData"))
load(here("data", "btp_cor_preg.RData"))

## ---- Anonymised plot data ---------------------------------------------------
load(here("data", "violin_out.RData"))

## ---- Exposure - outcome associations ----------------------------------------
load(here("data", "single_reg.RData"))

```

```{r, tidy-data}
time_labs <- c("Birth to 12 months", "Pregnancy")

## ---- Function to tidy descriptives ------------------------------------------
cleanData = function(x){

out <- x %>%
separate(
  data = .,
  col = variable, 
  sep = "_(?=[^_]*$)", 
  into = c("variable", "time")) %>%
mutate(
  .data = ., 
  time = factor(time, labels = time_labs, ordered = TRUE)) %>%
mutate(
  .data = ., 
  time = fct_relevel(time, rev(time_labs))) %>%
left_join(., names_neat, by = "cohort") %>%  
mutate(
  .data = .,
  cohort_neat = fct_relevel(
    cohort_neat, 
    c("MoBa", "DNBC", "GENR", "ALSPAC", "NINFEA", "EDEN Nancy", "EDEN Poitiers", 
      "INMA Gipuzkoa", "INMA Sabadell", "INMA Valencia"))) %>%
mutate(cohort_neat = factor(cohort_neat, ordered = TRUE))

}

desc_clean <- exposures.desc$continuous %>% cleanData

## ---- Violin plot data -------------------------------------------------------
violin.plotData <- violin_out %>% 
bind_rows(.id = "variable") %>% 
cleanData %>% 
group_by(variable, time, cohort) %>% 
mutate(n = n()) %>% 
mutate(cohort_n = paste0(cohort_neat, "\n", "n = ", n))


## ---- Exposure - outcome plot data -------------------------------------------
#single_reg.plotData <- single_reg.out %>% 
#left_join(., names_neat, by = "cohort") %>%  
#mutate(
#  .data = .,
#  cohort_neat = fct_relevel(
#    cohort_neat, 
#    rev(c("MoBa", "ALSPAC", "NINFEA", "Pooled results")))) %>%
#filter(variable != "intercept") %>%
#mutate(across(est:uppci, ~ exp(.x)))
```

\newpage

# Overview
I've put together (i) descriptive plots for exposures and outcomes, (ii) 
correlations between exposures, and (iii) associations between each exposure
and outcome. I will also look at non-linearity in exposure-outcome associations,
but I need to write a function to do spline modelling in DataSHIELD.

\newpage

# Descriptive statistics for exposures

## NDVI (300m buffer)
```{r ndvi, fig.showtext = TRUE, fig.height = 10, fig.width = 8}
hVPlot(var = "ndvi300", y_label = "NDVI", limits = c(0.1, 0.8)) 
```

\newpage

test text for new page
